Designing a high-performance boundary element library with OpenCL and Numba

### 2021-04-14 bem.tex
final changes

Once a triangulation is given we need to define the necessary data structures for the discretization. Bempp-cl uses a Galerkin discretization: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these function. We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N \dvec{x}_j\phi_j$, where $\dvec{x}$ is a vector of coefficients. In the most simple case, we can define the function \(\phi_j\) to be equal to 1 on the triangle $\tau_j$ and 0 everywhere else. Other spaces are commonly defined to be piecewise polynomials on each triangle.

@@

In Bempp-cl, an operator definition consists of the type of the operator (e.g. Laplace single-layer in the above example),
the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
consists of either the dual representation in the forms of the integrals $\dvec{b}_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx]$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.

Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$. For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_\ell$ and $\omega_q$ are the quadrature weights. In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see Figure \ref{fig:triangles}), corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.

>>>>

Once a triangulation is given we need to define the necessary data structures for the discretization. Bempp-cl uses a Galerkin discretization: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these functions. We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N \dvec{x}_j\phi_j$, where $\dvec{x}$ is a vector of coefficients. In the most simple case, we can define the function \(\phi_j\) to be equal to 1 on the triangle $\tau_j$ and 0 everywhere else. Other spaces are commonly defined to be piecewise polynomials on each triangle.

@@

In Bempp-cl, an operator definition consists of the type of the operator (e.g., Laplace single-layer in the above example),
the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
consists of either the dual representation in the forms of the integrals $\dvec{b}_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx]$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.

Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$. For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_\ell$ and $\omega_q$ are the quadrature weights. In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see \cref{fig:triangles}), corresponding singular quadrature rules are used that are based on singularity-removing coordinate transformations.

### 2021-04-14 bempp_overview.tex
final changes

The main user-visible component of Bempp-cl is the module \pyth{bempp.api}, which defines all user interface functions and other high-level routines. In particular, it contains the definition of the main object types: \pyth{Grid}, \pyth{Space}, \pyth{GridFunction}, \pyth{Operator}. The computational backend routines are contained in the module \pyth{bempp.core}. Currently, we support a Numba and OpenCL backends. An overview of this structure is provided in \cref{fig:overview}.

The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles. For larger problems this cost can be reduced through approximation techniques, such as hierarchical (H-) matrices or fast multipole methods (FMM) with log-linear or even linear complexity. The price of the improved complexity is significantly more involved data structures and additional approximation errors. Bempp-cl provides interfaces to Exafmm \cite{bempp_exafmm} for large problems. Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory. Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as highly as possible.

>>>>

The main user-visible component of Bempp-cl is the module \pyth{bempp.api}, which defines all user interface functions and other high-level routines. In particular, it contains the definitions of the main object types: \mbox{\pyth{Grid},} \mbox{\pyth{Space},} \mbox{\pyth{GridFunction},} and \pyth{Operator}. The computational backend routines are contained in the module \pyth{bempp.core}. Currently, we support Numba and OpenCL backends. An overview of this structure is provided in \cref{fig:overview}.

The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles. For larger problems, this cost can be reduced through approximation techniques, such as hierarchical (H-) matrices or fast multipole methods (FMM) with log-linear or even linear complexity. The price of the improved complexity is significantly more involved data structures and additional approximation errors. Bempp-cl provides interfaces to ExaFMM \cite{bempp_exafmm} for large problems. Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory. Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as highly as possible.

### 2021-04-14 intro.tex
final changes

\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations. The original release came with a simple Python wrapper to the C++ library \cite{bempp_orig}. Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl . This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use \cite{Bempp-cl}. Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp. In this article, we discuss the motivation for the rewrite and restructure of Bempp and the reasoning behind the design choices we made when writing Bempp-cl.

@@

Due to the three languages involved, making changes to the library would often mean having to duplicate changes in three places, with many class structures duplicated in all three languages. This made seemingly simple changes an onerous task, and provided a barrier to new members of the community looking to contribute to the open-source project.

>>>>

\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations. The original release came with a simple Python wrapper to the C++ library \cite{bempp_orig}. Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl. This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use \cite{Bempp-cl}. Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp. In this article, we discuss the motivation for the rewrite and restructure of Bempp and the reasoning behind the design choices we made when writing Bempp-cl.

@@

Due to the three languages involved, making changes to the library would often mean having to duplicate changes in three places, with many class structures duplicated in all three languages. This made seemingly simple changes into onerous tasks, and provided a barrier to new members of the community looking to contribute to the open-source project.

### 2021-04-14 main.tex
final changes

\begin{abstract}
The Bempp boundary element library is a well known library for the simulation of a range of electrostatic, acoustic and electromagnetic problems in homogeneous bounded and unbounded domains. It originally started as a traditional C++ library with a Python interface. Over the last two years we have completely redesigned Bempp as a native Python library, called Bempp-cl, that provides computational backends for OpenCL (using PyOpenCL) and Numba. The OpenCL backend implements kernels for GPUs and SIMD optimized for CPUs. In this paper we will discuss the design of Bempp-cl, provide performance comparisons on different compute devices and discuss the advantages and disadvantages of OpenCL as compared to Numba.
\end{abstract}

@@

\bibitem{bempp_orig}
W. \'{S}migaj, T. Betcke, S. Arridge, J. Phillips, and M. Schweiger,
``Solving boundary integral problems with BEM++,''
{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, Article 6, 2015.
\textsc{doi}: \href{https://doi.org/10.1145/2590830}{10.1145/2590830}

\bibitem{bempp_exafmm}
T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba, ``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-Exafmm''
{\it arXiv preprint} arXiv:103.01048, 2021.
\end{thebibliography}

>>>>

\begin{abstract}
The Bempp boundary element library is a well known library for the simulation of a range of electrostatic, acoustic and electromagnetic problems in homogeneous bounded and unbounded domains. It originally started as a traditional C++ library with a Python interface. Over the last two years we have completely redesigned Bempp as a native Python library, called Bempp-cl, that provides computational backends for OpenCL (using PyOpenCL) and Numba. The OpenCL backend implements kernels for GPUs and CPUs with SIMD optimization. In this paper, we discuss the design of Bempp-cl, provide performance comparisons on different compute devices, and discuss the advantages and disadvantages of OpenCL as compared to Numba.
\end{abstract}

@@

\bibitem{bempp_orig}
W. \'{S}migaj, T. Betcke, S. Arridge, J. Phillips, and M. Schweiger,
``Solving boundary integral problems with BEM++,''
{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, article 6, 2015.
\textsc{doi}: \href{https://doi.org/10.1145/2590830}{10.1145/2590830}

\bibitem{bempp_exafmm}
T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba,
``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-ExaFMM,''
2021.
arXiv (preprint): \href{https://arxiv.org/abs/2103.01048}{2103.01048}.
\end{thebibliography}

### 2021-04-14 opencl.tex
final changes

OpenCL splits computational tasks into work-items, each of which represents a single execution of a compute kernel. Work-items are grouped together into work-groups, which share local memory. Barrier synchronization is only allowed within a work-group. All work items are uniquely indexed by a one, two, or three dimensional index space, called NDRange. Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device. OpenCL allows kernels to be loaded as strings and compiled on-the-fly for a given device, making it well suited for launching from high-productivity languages.

@@

OpenCL has very good support for vectorized operations: it provides vector data types and defines a number of standard operations for these vector types. For example, the type \ocl{double4} will allow four double values to be held in a SIMD register. This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.

@@

\begin{itemize}
	\item Test and trial indices of triangles over which to be integrated.
	\item Signs of the normal directions for the spaces.
	\item Test and trial grid as flat floating point array, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
	\item Test and trial connectivity information that stores the indices of each node for each triangle.
	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
	\item Test and trial basis function multipliers for each triangle.
	\item Quadrature points and quadrature weights.
	\item A buffer that contains the global assembled matrix.
	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
	\item Number of test and trial degrees of freedom.
	\item A single byte that is set to one if the test and trial grids are disjoint.
\end{itemize}

@@

\subsection{SIMD optimized kernels}
When we are running on a CPU and want to take advantage of available SIMD optimizations, we need to make a few modifications to our approach. The corresponding kernel works similarly to what is described above, but we compute a batch of interactions between one test triangle and $X$ trial triangles, where $X$ is either 4, 8, or 16 (depending on the number of available SIMD lanes). This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation. If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard-non vectorized kernel.

@@

Alternatively, some CPU OpenCL runtime environments can (optionally) try and auto-vectorize kernels by batching together work items on SIMD lanes, similar to what we do manually. In our experience, this works well for very simple kernels but often fails for more complex OpenCL kernels. This is why we decided to implement this strategy manually.

>>>>

OpenCL splits computational tasks into work-items, each of which represents a single execution of a compute kernel. Work-items are grouped together into work-groups, which share local memory. Barrier synchronization is only allowed within a work-group. All work-items are uniquely indexed by a one, two, or three dimensional index space, called \ocl{NDRange}. Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device. OpenCL allows kernels to be loaded as strings and compiled on-the-fly for a given device, making it well suited for launching from high-productivity languages.

@@

OpenCL has very good support for vectorized operations: it provides vector data types and defines a number of standard operations for these vector types. For example, the type \mbox{\ocl{double4}} will allow four double values to be held in a SIMD register. This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.

@@

\begin{itemize}
	\item Test and trial indices of triangles over which to be integrated.
	\item Signs of the normal directions for the spaces.
	\item Test and trial grids as flat floating point arrays, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
	\item Test and trial connectivity information that stores the indices of each node for each triangle.
	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
	\item Test and trial basis function multipliers for each triangle.
	\item Quadrature points and quadrature weights.
	\item A buffer that contains the global assembled matrix.
	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
	\item The number of test and trial degrees of freedom.
	\item A single byte that is set to one if the test and trial grids are disjoint.
\end{itemize}

@@

\subsection{SIMD optimized kernels}
When we are running on a CPU and want to take advantage of available SIMD optimizations, we need to make a few modifications to our approach. The corresponding kernel works similarly to what is described above, but we compute a batch of interactions between one test triangle and $X$ trial triangles, where $X$ is either 4, 8, or 16 (depending on the number of available SIMD lanes). This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation. If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard non-vectorized kernel.

@@

Alternatively, some CPU OpenCL runtime environments can (optionally) try and auto-vectorize kernels by batching together work-items on SIMD lanes, similar to what we do manually. In our experience, this works well for very simple kernels but often fails for more complex OpenCL kernels. This is why we decided to implement this strategy manually.

### 2021-04-14 performance.tex
final changes

In this section, we provide a number of performance benchmarks. The tests are running on a Dell Precision 7740 Workstation Laptop with 64 GB RAM. Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz. The CPU supports AVX, AVX2, and AVX-512.
As a GPU we use an Nvidia Quadro RTX 3000 GPU. All benchmark tests were performed in Linux. For OpenCL on the GPU, we use the Nvidia GPU drivers and as CPU runtime we compare the open-source PoCL driver against the Intel CPU OpenCL runtime environment. All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba just-in-time compilers did not skew the results. Though hardly noticeable by users, for smaller experiments the compilation phase typically takes longer than the actual computation.

@@

We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the PoCL OpenCL runtime driver. We run the tests in single precision and double precision. The native vector width for both drivers in single precision is $8$, and in double precision is $4$, corresponding to AVX instructions. In Bempp-cl, this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision. Within the assembly, almost all floating point instructions are manually vectorized to take advantage of this. We should hence see up to a factor 2 speed-up between single and double precision assembly.

@@

\begin{figure}
\begin{center}
\begin{tabular}{l|c|c}
	        &   single      &    double\\
	        \hline
	 PoCL   &   0.05s       &    0.08s\\
	 Numba  &   0.25s       &    0.25s\\
	 GPU    &   0.11s       &    3.00s\\
\end{tabular}
\end{center}
\caption{Comparison of PoCL, Numba and GPU Assembly for a grid with 2048 elements.}
\label{fig:cpu_gpu_numba_compare}
\end{figure}
In Bempp-cl, we can easily switch between CPU Assembly, GPU Assembly, and Numba Assembly. \Cref{fig:cpu_gpu_numba_compare} shows a comparison between these modes for a grid with 2048 elements. The speed differences are striking. GPU assembly in single precision is a factor of two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision). Numba is five times slower for single precision and still around three times slower for double precision than PoCL. The GPU behaviour can be explained by data transfer: while the GPU kernels themselves are extremely fast, data transfer over the bus severely limits performance. Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with tens of thousands of rows and columns on the device.

@@

In this section, we consider the evaluation of the electric potential operator, defined by
\begin{multline}
\left(\mathcal{E}\mathbf{p}\right)(\bx) = \ii k\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\\
-\frac{1}{\ii k}\nabla_{\bx}\int_{\Gamma}\operatorname{div}\mathbf{p}(\by)g(\bx, \by)\ds[\by],
\end{multline}
where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\pi|\bx - \by|}$ is the Helmholtz Green's function, and
$k$ the wavenumber of the problem.
The function $\mathbf{p}:\mathbb{R}^3\to\mathbb{R}^3$ is a vector-valued function, leading to an overall vector solution at each point $\bx$. The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}. For these experiments, we again use a grid with 32,768 triangles but this time RWG (Rao, Wilton, and Glisson) edge-based basis functions. For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.

@@

\begin{figure}
	\center
	\includegraphics[width=7cm]{img/efield_domain_potential.pdf}
	\caption{Evaluation of an electric field potential operator on CPU via PoCL vs GPU}
	\label{fig:efield_domain_potential}
\end{figure}

\section{SUMMARY}
With Bempp-cl, we have created a Python library that achieves high-performance through use of modern just-in-time compilation technologies. Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines. Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.

Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends, allowing us to remain up to date with future compute models. OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU and GPU based compute devices with a simple parameter change. In this article, we demonstrate Nvidia benchmarks, the same benchmarking code could be used to run on AMD or Intel GPUs.

@@

The current focus of further developments is on letting Bempp-cl take advantage of cluster computing by integrating the \pyth{mpi4py} MPI bindings for Python. We have also made big steps forward for large problems by creating a black-box FMM interface that currently intenfaces to Exafmm: this has allowed us to solve problems with 10 million elements on a single workstation. We believe that Python-focused development (with some native routines in lower-level languages) is a scalable model and are aiming to exploit this scalability further as we move from single workstation computations to large cluster problems.

>>>>

In this section, we provide a number of performance benchmarks. The tests were all run on a Dell Precision 7740 Workstation Laptop with 64 GB RAM. Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz. The CPU supports AVX, AVX2, and AVX-512.
As a GPU we use an Nvidia Quadro RTX 3000 GPU. All benchmark tests were performed in Linux. For OpenCL on the GPU, we use the Nvidia GPU drivers and as CPU runtime we compare the open-source PoCL driver against the Intel CPU OpenCL runtime environment. All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba just-in-time compilers did not skew the results. Though hardly noticeable by users, for smaller experiments the compilation phase typically takes longer than the actual computation.

@@

We first compare the OpenCL CPU performance for the Intel OpenCL runtime and the PoCL OpenCL runtime driver. We run the tests in single precision and double precision. The native vector width for both drivers in single precision is $8$, and in double precision is $4$, corresponding to AVX instructions. In Bempp-cl, this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision. Within the assembly, almost all floating point instructions are manually vectorized to take advantage of this. We should hence see up to a factor 2 speed-up between single and double precision assembly.

@@

\begin{figure}
\begin{center}
\begin{tabular}{l|c|c}
	        &   single      &    double\\
	        \hline
	 PoCL   &   0.05s       &    0.08s\\
	 Numba  &   0.25s       &    0.25s\\
	 GPU    &   0.11s       &    3.00s\\
\end{tabular}
\end{center}
\caption{Comparison of PoCL, Numba and GPU assembly for a grid with 2048 elements.}
\label{fig:cpu_gpu_numba_compare}
\end{figure}
In Bempp-cl, we can easily switch between CPU Assembly, GPU Assembly, and Numba Assembly. \Cref{fig:cpu_gpu_numba_compare} shows a comparison between these modes for a grid with 2048 elements. The speed differences are striking. GPU assembly in single precision is a factor of two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision). Numba is five times slower for single precision and still around three times slower for double precision than PoCL. The GPU behaviour can be explained by data transfer: while the GPU kernels themselves are extremely fast, data transfer over the bus severely limits performance. Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with tens of thousands of rows and columns on the device.

@@

In this section, we consider the evaluation of the electric potential operator, defined by
\begin{multline}
\left(\mathcal{E}\mathbf{p}\right)(\bx) = \ii k\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\\
-\frac{1}{\ii k}\nabla_{\bx}\int_{\Gamma}\operatorname{div}\mathbf{p}(\by)g(\bx, \by)\ds[\by],
\end{multline}
where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\uppi|\bx - \by|}$ is the Helmholtz Green's function, and
$k$ the wavenumber of the problem.
The function $\mathbf{p}:\mathbb{R}^3\to\mathbb{R}^3$ is a vector-valued function, leading to an overall vector solution at each point $\bx$. The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}. For these experiments, we again use a grid with 32,768 triangles but this time RWG (Rao--Wilton--Glisson) edge-based basis functions are used. For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.

@@

\begin{figure}
	\center
	\includegraphics[width=7cm]{img/efield_domain_potential.pdf}
	\caption{Evaluation of an electric field potential operator on CPU via PoCL vs Nvidia GPU}
	\label{fig:efield_domain_potential}
\end{figure}

\section{SUMMARY}
With Bempp-cl, we have created a Python library that achieves high-performance through use of modern just-in-time compilation technologies. Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines. Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the ExaFMM library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.

Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends, allowing us to remain up to date with future compute models. OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU- and GPU-based compute devices with a simple parameter change. In this article, we demonstrate Nvidia benchmarks: the same benchmarking code could be used to run on AMD or Intel GPUs.

@@

The current focus of further developments is on letting Bempp-cl take advantage of cluster computing by integrating the \pyth{mpi4py} MPI bindings for Python. We have also made big steps forward for large problems by creating a black-box FMM interface that currently interfaces to ExaFMM: this has allowed us to solve problems with 10 million elements on a single workstation. We believe that Python-focused development (with some native routines in lower-level languages) is a scalable model and are aiming to exploit this scalability further as we move from single workstation computations to large cluster problems.
